---
title: "Chapter 4 and 5 notes"
author: "Anna Hedvig"
date: "9/23/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Exploring data with graphs

the role of the graph:
  -Show the data
  -induce the reader to think about the data being presneed (rather than some other aspect of the graph)
  -Avoid distorting the data
  -Present many numbers with minimum ink
  -Make larger data sets coherent
  -encourage the reader to compare different pieces of data
  - reveal data

Rules of thomb: x- and y-axis has informative labels, few distractions, and minimum ink, and using 2D-plots, and dont scale the y-axis to represent something false.

```{r}
installing ggplot2

library(ggplot2)
```

##the anatomy of a plot
the tree layers in the ggplot: each layer has a geometric element like bars, points, lines and text called a geom. the sizes and locations are determined by the aesthetic properties aes()

```{r}
Commonly used geoms:
  geom_bar() --> a variable to plot on the x-axis, creates a layer with bars showing statistical properties
  geom_point()--> layer showing the datapoints
  geom_line()--> layer that connetcs datapoints with a straight line
  geom_smooth() --> creates a layer with a smoother line...
  geom_histogram() -->hisogram on it
  goem_boxplot()--> box-whisker diagram
  geom_text()--> creates a layer of text
  geom_density() --> density plot
  geom_errorbar() --> creates an errorbar displayed on it
  geom_hline() -->user defined hori<ontal or vertical line, perspectivly
  
  in the () you type the aes() like x- and y-axis coordinates
```

##the anatomy of a ggplot() function

general version of a graph
```{r}
myGraph<-ggplot(myData, aes(variable for x-axis, variable for y-axis))

dataframe is myData. If one of the variables was gender, and we wanted that part coloured:

myGraph<-ggplot(myData, aes(variable for x-axis, variable for y-axis, colour= gender))
  
Adding layers to the myGraph

adding and poits (blue points): myGraph+geom_bar()+geom_point(colour="blue")

adding labels: myGraph+geom_bar()+geom_point(colour="blue")+labels(x="text", y="text")

wanting different types of variables (ex. ratings) to be different shapes:
  graph+geom_point(aes(shape=special ting))
```

##stats and geoms
the polits automatically calculates statistical properties when using geoms like geom_histogram() 
they are called built-in stats in ggplot2
 
see in p129 in the book

##avoiding overplotting
position="x", position adjustments
    - dodge: no overlap in the objects at the side
    - stack and fill: positions obetcs so they are stacked.
    - identitiy: no position adjustment
    - jitter: adds a random offset to objects so they dont overlap
    
Splitting a  plot into subgroups: producing a grid that splits the data displayed by the plot using facet_grid(), this is done by...? this is hard...

##regression and scatterplots
after plotting a scatterplot with the ex. myGraph+geom_point(), wanting to add a regression line, you add the geom geom_smooth(). the shaded part on the graph is the 95% confidence interval. if we want a straight line to represent the regression line, you add + geom_smooth(method ="lm") for linear model. if you want another colour: + geom_smooth(method ="lm", colour="red") 

switching the confidence interval off --> adding se = F to the geom_smooth like: geom_smooth(method="lm", se = F)

transparency and colour of confidence interval: geom_smooth(method="lm", alpha=0.1, fill="Blue")

##Grouped scatterplots
first seperating the groups in the scatterplot:
example with exam anxiety

```{r}
scatter<-ffplot(myData, aes(x-axis,y-axis,colour=variable, ex. gender))

after this we execute:

scatter+geom_point()+geom_smooth(method="lm",aes(fill=Variable ex. gender), alpha=0.1)

we fill, so the confidence intervals are calculated and shown in different colours

after this labels can be added: +labels(x="title", y="title", colour= variable ex. gender)

the title also differentiates in colour depending on the gender.
opts
  
```
 
##Histogams and problems with frequency distributions

outliners: a score very different from the rest, it can bias the mean

removing legend: +opts(legend.position="none"

Boxplot and box-whisker diagrams. center of plot is the median, sorrounded by the 50% top and 50% bottom of the observations. example of boxplot:

```{r}
festivalData<-read.delim("/Users/bruger/Desktop/ExpMeth-1---Assignment-1/Data_ex/DownloadFestival.dat",header=TRUE)

festivalHistogram<-ggplot(festivalData, aes(festivalData$day1))
festivalHistogram+geom_histogram()

```

```{r}
festivalHistogram+geom_histogram(binwidth = 0.4)+labs(x="Hygiene (day 1 of festival",y="Frequency")
```

Making a boxplot over genders:

```{r}
festivalBoxplot<-ggplot(festivalData, aes(gender, day1))
  festivalBoxplot+geom_boxplot()+labs(x="Gender", y="Hygiene(Day 1 of festival")

```

Taking away the outliners for woman and men:

```{r}

```


Density plots

```{r}
festivalData<-read.delim("/Users/bruger/Desktop/ExpMeth-1---Assignment-1/Data_ex/DownloadFestival(No Outlier).dat",header=TRUE)
density<-ggplot(festivalData, aes(day1))
density+geom_density()+labs(x="hygiene (day 1 of festival)", y= "Density Estimate")
```







